PhD Defense

### 2023-05-30 presentation/phd_defense_pres.tex
Populate PINODE part

@@

@@

\textbf{Objectives:}
\begin{itemize}
	\item Introduce $\mathcal{SR}3$ relaxation to linear mixed-effects models - $\mathcal{MSR}3$ (see \footcite{sholokhov2022relaxation}).
	\item Develop theoretical foundations for it (see \footcite{aravkin2022jimtheory}).
	\item Implement it as a \texttt{scikit-learn}-compatible Python package -- \texttt{pysr3} (see \footcite{sholokhov2023pysr3}).
\end{itemize}
\vspace{1em}

@@

\section{Experiments}
\subsection{Application to Synthetic Problems}

@@

\section{PINODE}

@@

1) People used to model physical systems with first-principle knowledge
2) Data-Driven modelling of dynamical systems became a big thing  
3) However, it requires a lot of data 
4) Incorporating prior knowledge is a big recent trend, so history does a spiral

@@

\begin{frame}{Incorporating Knowledge into Models}
1) There are multiple ways of incorporating knowledge into system 
4) The overall umbrella term for it is physics-informed machine learning
2) Some use the equations that model phenomena
3) Some take aspects of it, e.g. symmetries and preservation laws, and forces
A network to respect those
5) Our work falls into the first category of approaches
\end{frame}

\begin{frame}{Reduced-Order Models (ROMs)}
\includegraphics[width=\textwidth]{Figures/roms_1}
\end{frame}

\begin{frame}{Reduced-Order Models (ROMs)}
\includegraphics[width=\textwidth]{Figures/roms_2}
\end{frame}

\begin{frame}{Reduced-Order Models (ROMs)}
\includegraphics[width=\textwidth]{Figures/roms_3}
\end{frame}

\begin{frame}{Reduced-Order Models (ROMs)}
\includegraphics[width=\textwidth]{Figures/roms_4}
\end{frame}

\begin{frame}{Physics-Informed Loss}
%1) We introduce physics to this system by adding a term which regularizes latent gradient field. 
%2) In particular, it forces it to be equal to what a true physics should be under such projection.
%3) We can not evaluate physics everywhere but we can at particular carefully-selected points. We call these points collocation points.
%4) We feed a lot of collocation points and ask a network to do interpolation 
\includegraphics[width=\textwidth]{Figures/collocations_1.png}
\end{frame}

\begin{frame}{Physics-Informed Loss}
	\includegraphics[width=\textwidth]{Figures/collocations_2.png}
\end{frame}

\begin{frame}{Physics-Informed Loss}
	\includegraphics[width=\textwidth]{Figures/collocations_3.png}
\end{frame}

\begin{frame}{Physics-Informed Loss}
	\includegraphics[width=\textwidth]{Figures/collocations_4.png}
\end{frame}

\begin{frame}{Physics-Informed Loss}
	\includegraphics[width=\textwidth]{Figures/collocations_5.png}
\end{frame}

@@

Projection to a high-dimensional (128) space:
\begin{equation}
    \label{eq:duffing_true_decoder}
    \bd{x} := \mathcal{A}(\bd{z}) = A\bd{z}^3, \quad A \in \mathbb{R}^{128 \times 2}, \quad A_{ij} \sim_{i.i.d.} \mathcal{N}(0, 1)
\end{equation}

\begin{figure}
	\centering
	\includegraphics[width=0.8\textwidth]{Figures/duff_results.png}
\end{figure}

@@

\begin{frame}{Results: Stable Long-Term Predictions}
\begin{figure}
	\begin{subfigure}[b]{0.3\textwidth}
		\includegraphics[width=\textwidth]{Figures/duff_full_train.pdf}
	\end{subfigure}%
	\begin{subfigure}[b]{0.7\textwidth}
		\centering
		\includegraphics[width=0.7\textwidth]{Figures/duffing_periods.pdf}
	\end{subfigure}
\end{figure}
\begin{figure}
	\includegraphics[width=0.7\textwidth]{Figures/duffing_periods_examples.png}
\end{figure}
\end{frame}

\begin{frame}{Results: Learning From Collocations}
1) Finally we show that collocations can be even more useful than the data itself.
2) The difference is especially prominent in low-data regime.
3) It shows that collocations are powerful source of information and that the network can indeed interpolate between them.
\end{frame}

@@

\appendix
\begin{frame}{Designing an Algorithm}
	$G_{\nu,\eta}$ encodes both gradient of a Lagrangian (lines 1-2) and the complementarity condition (line 3):
	\eq{
		G_{\nu,\eta}((\beta,\gamma,v),(\tbeta,\tgamma)) := \begin{bmatrix}
			\nabla_\beta \LL(\beta, \gamma) + \eta(\beta-\tbeta) \\
			\nabla_\gamma \LL(\beta, \gamma) + \eta(\gamma-\tgamma) - v\\
			v \bigodot \gamma - \mu\textbf{1}
		\end{bmatrix}
	}
	We apply Newton method to $G$ while geometrically decreasing $\mu$. \\
\textbf{Lemma:} For every $(\mu,\eta) \in \R_+\times\R_{++}$,
\eq{
	&(\hat\beta,\hat\gamma) = \argmin_{(\beta,\gamma)}\LL_{\eta,\mu}((\beta, \gamma),(\tbeta, \tgamma)) \\
	& \iff \\
	& \exists \hat{v} \in \R_{+}^q \text{ s.t. } G_{\nu,\eta}((\beta,\gamma,\hat{v}),(\tbeta,\tgamma)) = 0
}
If $\mu > 0$, then $\hat{v} = -\nabla\phi_\mu(\hat{\gamma})$, and if $\mu = 0$, then $\hat{v}$ is the unique KKT multiplier associated with the constraint $0 \leq \gamma$.
\end{frame}

@@

\begin{frame}{References}
\printbibliography
\end{frame}

>>>>

\AtBeginSection[]{
  \begin{frame}
  \vfill
  \centering
  \begin{beamercolorbox}[sep=8pt,center,shadow=true,rounded=true]{title}
    \usebeamerfont{title}\insertsectionhead\par%
  \end{beamercolorbox}
  \vfill
  \end{frame}
}

@@

\section{$\mathcal{MSR}3$ -- Sparse Relaxed Regularized Regression for Linear Mixed-Effect Models}

@@

\textbf{Objectives:}
\begin{itemize}
	\item Extend $\mathcal{SR}3$ relaxation to linear mixed-effects models - $\mathcal{MSR}3$ (see \footcite{sholokhov2022relaxation}).
	\item Develop theoretical foundations for it (see \footcite{aravkin2022jimtheory}).
	\item Implement it as a \texttt{scikit-learn}-compatible Python package -- \texttt{pysr3} (see \footcite{sholokhov2023pysr3}).
\end{itemize}
\vspace{1em}

@@

@@

\section{Physics-Informed Neural ODE (PINODE): Embedding Physics into Models using Collocation Points}

@@

%1) People used to model physical systems with first-principle knowledge
%2) Data-Driven modelling of dynamical systems became a big thing  
%3) However, it requires a lot of data 
%4) Incorporating prior knowledge is a big recent trend, so history does a spiral

\uncover<2->{
\begin{columns}[T,onlytextwidth]
\column{0.6\textwidth}
\textcolor{orange}{First-Principle Models:}
\begin{itemize}
	\item Require extensive knowledge of the phenomenon
	\item Require a lot of compute for simulating large-scale phenomena
\end{itemize}
\column{0.4\textwidth}
\centering
\includegraphics[width=\textwidth]{Figures/NN.jpg}
\end{columns}
}
\vspace{2em}
\uncover<3->{
\begin{columns}[T,onlytextwidth]
\column{0.6\textwidth}
\textcolor{green}{Data-Driven Models:}
\begin{itemize}
	\item Require a lot of data
	\item Often struggle to extrapolate
\end{itemize}
\column{0.4\textwidth}
\centering
\includegraphics[width=0.8\textwidth]{Figures/nn_schematics.pdf}
\end{columns}
}
\vspace{2em}
\uncover<4->{
\begin{columns}[T,onlytextwidth]
\column{0.6\textwidth}
\textcolor{blue}{Hybrid Models: }
\begin{itemize}
	\item Incorporate elements of both approaches
	\item Supplement data with knowledge or priors
\end{itemize}
\column{0.4\textwidth}
\centering
\includegraphics[width=0.7\textwidth]{Figures/NS_pic.png}
\end{columns}
}
\end{frame}

\begin{frame}{Incorporating Knowledge of Physics into Neural Networks}
\begin{enumerate}
	\item<2-> \textbf{Symmetry Based:} incorporate symmetries and conservation laws as hard constraints into a network: E3NNs\footcite{geiger2022e3nn}, LieConv\footcite{Finzi2020}, RiCNN \footcite{chidester2018rotation}
	\item<3-> \textbf{Model Discovery Based:} use a library of terms to discover simple equations: SINDy\footcite{champion2019data}, Genetic Programming\footcite{schmidt2009distilling}
	\item<4-> \textbf{Equations Based:} incorporate first-principle models to aid training of networks: PINNs\footcite{raissi2017physics}, UDEs\footcite{rackauckas2020udes}
\end{enumerate}

\begin{columns}[T,onlytextwidth]
\column{0.3\textwidth}
	\uncover<2->{\includegraphics[width=0.7\textwidth]{Figures/ricnn.png}}
\column{0.7\textwidth}
	\uncover<3->{\includegraphics[width=\textwidth]{Figures/sindy.png}}
\end{columns}

@@

\begin{frame}{Reduced-Order Models (ROMs)}
\only<1>{\includegraphics[width=\textwidth]{Figures/roms_1}}
\only<2>{\includegraphics[width=\textwidth]{Figures/roms_2}}
\only<3>{\includegraphics[width=\textwidth]{Figures/roms_3}}
\only<4>{\includegraphics[width=\textwidth]{Figures/roms_4}}
\end{frame}

\begin{frame}{Physics-Informed Loss}
%1) We introduce physics to this system by adding a term which regularizes latent gradient field. 
%2) In particular, it forces it to be equal to what a true physics should be under such projection.
%3) We can not evaluate physics everywhere but we can at particular carefully-selected points. We call these points collocation points.
%4) We feed a lot of collocation points and ask a network to do interpolation 
Using chain rule:
\only<1>{\includegraphics[width=\textwidth]{Figures/collocations_1.png}}
\only<2>{\includegraphics[width=\textwidth]{Figures/collocations_2.png}}
\only<3>{\includegraphics[width=\textwidth]{Figures/collocations_3.png}}
\only<4>{\includegraphics[width=\textwidth]{Figures/collocations_4.png}}
\only<5>{\includegraphics[width=\textwidth]{Figures/collocations_5.png}}
\end{frame}

@@

Projection to a high-dimensional (128) space:
\begin{equation}
    \label{eq:duffing_true_decoder}
    \bd{x} := \mathcal{A}(\bd{z}) = A\bd{z}^3, \quad A \in \mathbb{R}^{128 \times 2}, \quad A_{ij} \sim_{i.i.d.} \mathcal{N}(0, 1)
\end{equation}
\uncover<2->{
\begin{figure}
	\centering
	\includegraphics[width=0.8\textwidth]{Figures/duff_results.png}
\end{figure}
}
\end{frame}

\begin{frame}{Results: Stable Long-Term Predictions}
\begin{figure}
	\begin{subfigure}[b]{0.3\textwidth}
		\includegraphics[width=\textwidth]{Figures/duff_full_train.pdf}
	\end{subfigure}%
	\begin{subfigure}[b]{0.7\textwidth}
		\centering
		\includegraphics[width=0.7\textwidth]{Figures/duffing_periods.pdf}
	\end{subfigure}
\end{figure}
\uncover<2->{
\begin{figure}
	\includegraphics[width=0.7\textwidth]{Figures/duffing_periods_examples.png}
\end{figure}
}
\end{frame}

\begin{frame}{Results: Burgers' Equation}
Burgers' Equation:
\begin{equation}
	\begin{split}
		& u_t  + uu_x = \nu u_{xx} \\
	    & u(-\pi, t) = u(\pi, t),\quad \forall t \in [0, T]
	\end{split}
\end{equation}
\uncover<2->{
\includegraphics[width=\textwidth]{Figures/burgers_examples_of_ics.pdf}
}
\uncover<3->{
\includegraphics[width=\textwidth]{Figures/example_burgers.pdf}
}

@@

\begin{frame}{Results: Learning From Collocations}
\includegraphics[width=\textwidth]{Figures/burgers_examples_of_ics.pdf}
\includegraphics[width=\textwidth]{Figures/data_vs_collocations.pdf}
\end{frame}

\begin{frame}{Discussion and Limitations}
We showed that:
\begin{itemize}
	\item Physics-informed loss improves accuracy and forecasting stability of ROMs
	\item Collocations can supplement data to improve model's performance for unseen initial conditions.
\end{itemize}
\vspace{2em}
Limitations:
\begin{itemize}
	\item Optimal choice of collocations is problem-specific
	\item Need a lot of collocations
	\begin{itemize}
		\item \textit{Could} be possible to overcome with smarter sampling techniques
	\end{itemize}
\end{itemize}
\end{frame}

\section{Single pixel imaging of spatio-temporal flows using differentiable latent dynamics}

@@

\appendix
\section{Appendix: $\mathcal{MSR}3$}
\begin{frame}{Designing an Algorithm}
	$G_{\nu,\eta}$ encodes both gradient of a Lagrangian (lines 1-2) and the complementarity condition (line 3):
	\eq{
		G_{\nu,\eta}((\beta,\gamma,v),(\tbeta,\tgamma)) := \begin{bmatrix}
			\nabla_\beta \LL(\beta, \gamma) + \eta(\beta-\tbeta) \\
			\nabla_\gamma \LL(\beta, \gamma) + \eta(\gamma-\tgamma) - v\\
			v \bigodot \gamma - \mu\textbf{1}
		\end{bmatrix}
	}
	We apply Newton method to $G$ while geometrically decreasing $\mu$. \\
\textbf{Lemma:} For every $(\mu,\eta) \in \R_+\times\R_{++}$,
\eq{
	&(\hat\beta,\hat\gamma) = \argmin_{(\beta,\gamma)}\LL_{\eta,\mu}((\beta, \gamma),(\tbeta, \tgamma)) \\
	& \iff \\
	& \exists \hat{v} \in \R_{+}^q \text{ s.t. } G_{\nu,\eta}((\beta,\gamma,\hat{v}),(\tbeta,\tgamma)) = 0
}
If $\mu > 0$, then $\hat{v} = -\nabla\phi_\mu(\hat{\gamma})$, and if $\mu = 0$, then $\hat{v}$ is the unique KKT multiplier associated with the constraint $0 \leq \gamma$.
\end{frame}

@@

\begin{frame}[noframenumbering,plain,allowframebreaks]{References}
    \printbibliography[heading=none]
\end{frame}
